Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-02-22 ms.tex
omg

This paper is organized as followed.  In $\S\ref{sec:model}$ we
describe our model, including the sample of galaxies used in our
analysis ($\S\ref{sec:gal_model}$) and our numerical procedure for
calculating the steady-state hydrodynamic profile of the CNM
($\S\ref{sec:hydro}$).  In $\S\ref{sec:results}$ we describe our
results.  In $\S\ref{sec:applications}$ we describe applicaiton of our
results, including a comparison of our gas profiles to those inferred
from {\it Chandra} observations and to the measured $L_X-\Mbh$
relationship.  In $\S\ref{sec:discussion}$ we discuss our results.  In
$\S\ref{sec:summary}$ we summarize our results.  Table \ref{table:definitions} provides the definitions of commonly used variables.

@@

\section{Applications}
\label{sec:applications}

@@

Figure \ref{fig:vwSources} shows that the average heating rate is relatively constant with SMBH mass.  For fixed wind heating $\dot{M}/\dot{M}_{\rm edd}$ increases with SMBH mass $\propto M_{\bullet}^{0.5(0.8)}$ for core(cusp) galaxies, respectively (Fig.~\ref{fig:mdot_mass}; eq.~[\ref{eq:eddr_analytic}]).  This appears in contradiction with the nuclear X-ray luminosities of quiescent galaxies, which though showing broad scatter instead obey $L_X \sim \Mbh^\alpha$, where $\alpha\simeq 0.7-0.8$ \citep{Pellegrini10, Miller+15}.  Our model instead predicts that $\Mdot
\sim \Mbh^{1.5-2}$, such that if $L_x\sim M$, then we would have $\L_x
\sim \Mbh^{1.5-2}$, steeper than the observed relationship.

This is shown Figure~\ref{fig:miller} shows our predicted ${\rm
  L_{X}/L_{\rm edd}=\epsilon_X \Mdot c^2}$ for core (blue) and cusp
  (green) galaxies superimposed with x-ray measurements (red stars),
  upper limits (red triangles) from \citet{Miller+15}. $\epsilon_X$ is
  an efficiency factor and $\Mdot$ is calculated using
  Equation~\eqref{eq:mdot_analytic}, with parameters consistent with
  the average star formation history for each $\Mbh$ (see the black
  lines in Figures~\ref{fig:vwSources} and~\ref{fig:eta}). Our
  predicted trend in  $L_{X}/L_{\rm edd}$ is steeper than the 
  trend found in \citet{Miller+15} (black line). Note that we try two 
  different prescriptions for $\epsilon_X$: $\epsilon_X=10^{-4}$ in
  the left-hand panel and $\epsilon_X=10^{-1} \Mdot/0.01 \MdotEdd$
  right-hand panel. In either case our predicted trend in $
  L_{X}/L_{\rm edd}$ is opposite the observationally inferred
  one. {\bf AG-Need a bit more discussion here.} We note that in the
  right hand panel the observed galaxies are below the upper limits
  from Ia supernovae and thermal instability and so may be accreting
  in a thermally stable state {\bf AG-but it does not mean that they
    actually are}.

@@

\subsection{Comparison to Observations}
{\bf AG: I do want some discussion of reproducing the density profiles
for individual observations, but I am not sure how to do it with our
new gridded approach.  BDM: too bad!}
% \subsubsection{Comparison to measured temperature and density profiles}
% \citet{AllenDunn+:2006a} use {\it Chandra} X-ray observations to infer the radial temperature and density profiles in the nuclei of nine nearby X-ray luminous elliptical galaxies.  \citet{RussellMcNamara+:2013a} performed a similar analysis with a somewhat enlarged galaxy sample.  Interestingly, these works find a correlation between the accretion rate inferred assuming Bondi accretion and the AGN jet power, which is estimated by the PdV work required to inflate observed cavities in the X-ray emitting gas.  
% % At some point it would be useful to describe the differences between
% % these two sets of observations \citealt{RussellMcNamara+:2013a}
% % tries several different extrapolation schemes for the gas density,
% % and finds that that the uncertainty in extrapolating the density
% % profile is the dominant source of uncertainty in measuring the
% % accretion rate. Find total number which overlaps once the Russell
% % galaxies are included...

@@

@@

\subsection{Cooling}
\label{sec:cooling}

\section{Discussion}

  \section{Summary}
  \label{sec:summary}
  We have calculated steady-state models for circumnuclear medium of a
  sample of early type galaxies, assuming that gas is supplied
  entirely by stellar wind mass loss and heated by shocked stellar
  winds and SNe Ia. We may use these profiles to calculate an $\Mdot$
  onto each of sample galaxy's SMBH, and get the scaling of $\Mdot$
  with $\Mbh$. We may then compare our results with observed trends of
  nuclear x-ray luminosity. Our main conclusions may be summarized as
  follows.

>>>>

This paper is organized as followed.  In $\S\ref{sec:model}$ we
describe our model, including the sample of galaxies used in our
analysis ($\S\ref{sec:gal_model}$) and our numerical procedure for
calculating the steady-state hydrodynamic profile of the CNM
($\S\ref{sec:hydro}$).  In $\S\ref{sec:results}$ we describe some
analytic results, while in $\S\ref{sec:numerical}$ we describe our numerical results.  In $\S\ref{sec:discussion}$ we discuss the implications of our results, to topics including...   In $\S\ref{sec:conclusions}$ we summarize our conclusions.  Table \ref{table:definitions} provides the definitions of commonly used variables.  Appendix \ref{app:rs} provides useful analytic results, while Appendix \ref{app:windheat} describes our method for calculating stellar wind heating and mass input in detail.

@@

@@

Figure \ref{fig:vwSources} shows that the CNM heating rate is on average approximately constant with SMBH mass, except for the largest black holes with $M_{\bullet} \gtrsim 5\times 10^{8}M_{\odot}$.  For fixed wind heating $\dot{M}/\dot{M}_{\rm edd}$ increases with SMBH mass $\propto M_{\bullet}^{0.5(0.8)}$ for core(cusp) galaxies, respectively (Fig.~\ref{fig:mdot_mass}; eq.~[\ref{eq:eddr_analytic}]).  This appears in contradiction with the nuclear X-ray luminosities of quiescent galaxies, which (though showing broad scatter) instead on average obey $\langle L_X/L_{\rm edd}\rangle \sim \Mbh^\alpha$, where $\alpha\simeq -0.25$ \citep{Pellegrini10, Miller+15}.  Figure~\ref{fig:miller} shows our predicted value of $\langle L_{X}/L_{\rm edd} \rangle =\epsilon_X \Mdot c^2$ as a function of black hole mass for core (blue) and cusp (green) galaxies.  The left panel assumes a constant low radiative efficiency $\epsilon_X = 10^{-4}$ typical of those  estimated for low luminosity AGN (e.g., \citealt{DiMatteo+00}, \citealt{Ho:2009a}) and SgrA* (e.g., \citealt{Quataert+99}, \citealt{Quataert:2004a}), while the right panel shows an efficiency of $\epsilon_X= \dot{M}/\dot{M}_{\rm edd}$ that depends on the Eddington ratio in the way predicted by simple models for radiatively inefficient accretion flows (\citealt{Narayan&Yi95}).  Superimposed with the theoretical curves are symbols showing X-ray measurements (red stars) and upper limits (red triangles) from \citet{Miller+15}, along with the best power-law fit (red line; {\bf BDM: change to this!}) and the approximate X-ray detection threshold (gray line).  Finally, shown with purple and orange lines are the maximum accretion rates for thermally stable accretion (eq.~[\ref{eq:Mdotmax}]) and set by Ia blow-out (eq.~[\ref{eq:eddr_Ia}]), as well as the maximum accretion rate allowed by Compton heating (eq.~[\ref{eq:MdotC}]).

For either of our assumptions about radiative efficiency our predicted trend in $\langle L_{X}/L_{\rm edd}\rangle$ for core and cusp galaxies is opposite the observationally inferred one.  However, the observed $\langle L_{X}/L_{\rm edd}\rangle$ do imply the accretion rates of observed galaxies are below the maximum ones for thermally stable accretion if $\epsilon_X \lesssim 10^{-4}$.  This suggests that most nearby galaxies are consistent with accreting in a thermally stable steady state, even if our estimate of their predicted heating does not match the observed trend with $M_{\bullet}$ perfectly.  On the other hand if $\epsilon_{X} \sim 10^{-5}$ then the trend would be consistent with the accretion rate set by Ia blow-out.  Also note that in the case of $\dot{M}-$dependent radiative efficiency ($\epsilon_X = \dot{M}/\dot{M}_{\rm edd}$), the predicted branches of  $\langle L_{X}/L_{\rm edd}\rangle$ decreases with decreasing $M_{\bullet}$, calling into question attempts to extrapolate the $\langle L_{X} \rangle - M_{\bullet}$ relationship to low $M_{\bullet}$ in order to constrain the occupation fraction of black holes in the lowest mass galaxies (\citealt{Miller+15}).

@@

% \subsubsection{Comparison to measured temperature and density profiles}
% \citet{AllenDunn+:2006a} use {\it Chandra} X-ray observations to infer the radial temperature and density profiles in the nuclei of nine nearby X-ray luminous elliptical galaxies.  \citet{RussellMcNamara+:2013a} performed a similar analysis with a somewhat enlarged galaxy sample.  Interestingly, these works find a correlation between the accretion rate inferred assuming Bondi accretion and the AGN jet power, which is estimated by the PdV work required to inflate observed cavities in the X-ray emitting gas.  
% % At some point it would be useful to describe the differences between
% % these two sets of observations \citealt{RussellMcNamara+:2013a}
% % tries several different extrapolation schemes for the gas density,
% % and finds that that the uncertainty in extrapolating the density
% % profile is the dominant source of uncertainty in measuring the
% % accretion rate. Find total number which overlaps once the Russell
% % galaxies are included...

@@

\section{Conclusions}
\label{sec:conclusions}
  We have calculated steady-state models for circumnuclear medium of a
  sample of early type galaxies, assuming that gas is supplied
  entirely by stellar wind mass loss and heated by shocked stellar
  winds and SNe Ia. We may use these profiles to calculate an $\Mdot$
  onto each of sample galaxy's SMBH, and get the scaling of $\Mdot$
  with $\Mbh$. We may then compare our results with observed trends of
  nuclear x-ray luminosity. Our main conclusions may be summarized as
  follows.

\section{Death Row}

@@
